manuscript_figures

### 2024-07-10 scripts/manuscript_figures.Rmd
added functional group change over time data

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(vegan)
library(RColorBrewer)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(gridExtra)
library(grid)

@@

# Calculate habitat productivity index for all transects

@@

# Plot three plots together using patchwork for production and HPI for fish and kelp biomass
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.nonlog <- production_plot | hpi.plot.nonlog +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog

@@

# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.log <- production_plot | hpi_plot_log +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.log

@@

>>>>

# Load libraries 
```{r}
library(plyr)
library(dunn.test)
library(kableExtra)
library(vegan)
library(RColorBrewer)
library(flextable)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(gridExtra)
library(grid)

@@

# Calculate habitat productivity index for all transects using HPI = average kelp biomass/area surveyed (80)
# Define the area surveyed (80 square meters for each transects)
area_surveyed <- 80

# Calculate the HPI for each transect
HPI.kelp <- HPI.kelp %>%
  mutate(HPI = total_kelp_biomass / area_surveyed)

# Summarize mean and standard error of HPI by treatment
HPI_summary <- HPI.kelp %>%
  group_by(TREATMENT) %>%
  summarise(
    mean_HPI = mean(HPI, na.rm = TRUE),
    se_HPI = sd(HPI, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  )

# Create the box plot
hpi_box_plot <- ggplot(HPI.kelp, aes(x = TREATMENT, y = HPI, fill = TREATMENT)) +
  geom_boxplot() +
  geom_point(data = HPI_summary, aes(x = TREATMENT, y = mean_HPI), shape = 21, size = 3, position = position_dodge(0.75)) +
  labs(x = "Treatment", y = "Habitat Productivity Index (HPI)") +  # Rename legend text
  scale_fill_manual(values = c("Control" = "#4E79A7", "Kelp" = "#F28E2B")) +  # Manual fill colors and labels
  theme_bw() +
  theme(axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18));hpi_box_plot

@@

# Plot three plots together using patchwork for production and HPI for fish and kelp biomass
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.nonlog <- (production_plot | hpi.plot.nonlog | hpi_box_plot) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.nonlog

@@

# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi.log <- (production_plot | hpi_plot_log | hpi_box_plot) +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi.log

@@

# Figure for functional group change in abundance over time using invert count data
```{r}
# Check for normality of residuals
# Log transform the abundance data
invert.count$log_abundance <- log(invert.count$abundance + 1.1)

# Check for normality of transformed data
anova_model <- aov(log_abundance ~ type * YEAR, data = invert.count)
residuals <- residuals(anova_model)

# Plot Q-Q plot
qqnorm(residuals)
qqline(residuals, col = "blue")

# Perform Shapiro-Wilk test
shapiro_test <- shapiro.test(residuals)
shapiro_test

# still not normal, use non-parametric test

# Perform Kruskal-Wallis test for interaction type and YEAR
interaction_data <- invert.count %>%
  mutate(interaction = interaction(type, YEAR))

kruskal_test_interaction <- kruskal.test(abundance ~ interaction, data = interaction_data)
print(kruskal_test_interaction)

# Extract relevant information from the test result
kruskal_summary <- data.frame(
  Test = "Kruskal-Wallis",
  Interaction = as.character(kruskal_test_interaction$method),
  ChiSquare = kruskal_test_interaction$statistic,
  DF = kruskal_test_interaction$parameter,
  P.Value = kruskal_test_interaction$p.value
)

# Create flextable object
flex <- flextable(kruskal_summary)

# Save flextable as PNG
png_file <- "output/model.result.tables/kruskal_test_interaction_summary.png"
save_as_image(flex, png_file)

# Assuming invert.count is your dataframe and abundance_g_per_m2 is already calculated
total_abundance <- invert.count %>%
  group_by(YEAR) %>%
  summarise(total_abundance_g_per_m2 = sum(abundance))

# merge total abundance with original data
invert.abun <- merge(invert.count, total_abundance, by = "YEAR")

# calculate relative abundance
invert.abun <- invert.abun %>%
  mutate(relative_abundance = (abundance / total_abundance_g_per_m2) * 100)

# Calculate mean and standard error per functional group and year
mean_std_data <- invert.abun %>%
  group_by(YEAR, type) %>%
  summarise(
    mean_rel_abundance = mean(relative_abundance),
    std_error = sd(relative_abundance) / sqrt(n()))

# Convert YEAR to factor
mean_std_data$YEAR <- factor(mean_std_data$YEAR)

# Example: Renaming groups in mean_std_data dataframe
mean_std_data <- mean_std_data %>%
  mutate(type = case_when(
    type == "crustacean" ~ "Crustaceans",
    type == "echinoderms" ~ "Echinoderms",
    type == "molluscs" ~ "Molluscs",
    TRUE ~ type  # Keep other cases unchanged
  ))

# Plot mean abundance with error bars
func.group.plot.line <- ggplot(mean_std_data, aes(x = YEAR, y = mean_rel_abundance, color = type, group = type)) +
  geom_errorbar(aes(ymin = mean_rel_abundance - std_error, ymax = mean_rel_abundance + std_error), width = 0.2) +
  geom_line() +
  geom_point(size = 3) +
  labs(x = "Year", y = "Relative Abundance (%)", color = "Functional Group") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18));func.group.plot.line

# Save the plot to a file
ggsave("output/figures/functional.plot.line.png", plot = func.group.plot.line, width = 15, height = 10)

# Plot stacked bar chart
stacked_bar_plot <- ggplot(mean_std_data, aes(x = factor(YEAR), y = mean_rel_abundance, fill = type)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  labs(x = "Year", y = "Relative Abundance (%)", fill = "Functional Group") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  scale_y_continuous(limits = c(0, 50), expand = c(0, 0));stacked_bar_plot

# Save the plot to a file
ggsave("output/figures/functional.plot.stacked.png", plot = stacked_bar_plot, width = 15, height = 10)

```
	Kruskal-Wallis rank sum test

data:  abundance by interaction
Kruskal-Wallis chi-squared = 104.02, df = 23, p-value = 2.83e-12
